\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2018-10-01 Documents/M1/Astronum_2018/sections/Conclusion.tex
Added conclusion section.

This is the conclusion section ...

>>>>

We have developed IMEX schemes for the two-moment neutrino transport in \texttt{thornado} code respecting Fermi-Dirac statistics.
The schemes employ algebraic closure based on Fermi-Dirac statistics, a first-order discontinuous Galerkin method, the simple Lax-Friedrichs flux with the convex-invariant time integrator to maintain point-wise realizability of the moments.
Since the realizability-preserving property is obtained from convexity of the realizable domain, it's possible to construct a scheme with high order DG method.
See \cite{chu_2018} for some examples with third-order DG method employed.

In the applications motivating this work, the neutrino distribution function can vary from 0 to 1, and we have considered the algebraic closure respecting Fermi-Dirac statistics for both low and high occupied condition.
Different values given by different algebraic closures were compared with the boundary values given by the Fermi-Dirac statistics at different occupation.
Among the algebraic closures we considered, Kershaw\cite{kershaw_1976}, Wilson\cite{wilson1975,leblanc1970}, Levermore\cite{levermore_1984}, Minerbo \cite{minerbo_1978}, Janka 1\cite{janka1991} Janka 2\cite{janka1992} and Cernohorsky \& Bludman \cite{cernohorskyBludman_1994}, only Cernohorsky \& Bludman closure respecting Fermi-Dirac statistics for any occupancy.
And it is Cernohorsky \& Bludman closure that we employed in the numerical tests in Section~\ref{se:NumericalTests}.

Following the work we have done in \cite{chu_2018}, two PD-ARSs have been proposed.
The one with SSPRK2 is second-order and strong stability-preserving in the streaming limit while other with SSPRK3 is third-order.
Their properties, accuracy and convex-invariant, are demonstrated with numerical tests.

In this work, we adopted Cartesian coordinates, linear collision term, non-relativistic, and fixed material background.
More realistic problems of scientific interest, such as with the scattering with energy changes and relativistic effects, are left for future research.

### 2018-10-01 Documents/M1/Astronum_2018/sections/Numerical.tex
Added conclusion section.

\subsubsection{Sine Wave Damping}
Figure~\ref{fig:SineWaveDamping} shows the numerical result of sine wave damping test.
See \cite{chu_2018} for more information about the test setting.
SSP2332 as a second-order scheme display second order convergence rate.
PD-ARSs convergent in first order, as a price paid for realizability-preserving.
\begin{figure}[h]
  \centering
    \includegraphics[width=0.7\textwidth]{figures/SineWaveDamping}
   \caption{Relative error versus number of elements $N$ for the damping sine wave test. Results for different values of the absorption opacity $\sigma_{\Ab}$, employing various IMEX time stepping schemes, are compared.  Errors for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are SSP2332 ($+$), PD-ARS with SSPRK2 (circles) and PD-ARS with SSPRK3 (hexagram).  Black dash-dot reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively.}
  \label{fig:SineWaveDamping}
\end{figure}

\subsubsection{Sine Wave Diffusion}
The last test with known smooth solution is sine wave diffusion test.
See \cite{chu_2018} for more information about the test setting.
Figure~\ref{fig:SineWaveDiffusionJ} shows the numerical result.
SSP2332 and PD-ARSs display third-order accuracy for the number density $\cJ$ and second-oder accuracy for $\cH_{x}$ and their errors are difficult to distinguish.
PD-ARS with SSP2332 behaviors as good as SSP2332 in the diffusion region but requires 1/3 less implicit solver.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionJ}
    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionH}
  \end{tabular}
   \caption{Absolute error for the number density $\cJ$ (left) and the number flux $\cH_{x}$ (right) versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are:  SSP2332 ($+$), PD-ARS with SSPRK2 (circles) and PD-ARS with SSPRK3 (hexagram). Black dash-dot on the left plot reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively. Black dash-dot on the right plot reference lines are proportional to $N^{-2}$ .}
   \label{fig:SineWaveDiffusionJ}
\end{figure}

>>>>

\subsubsection{Sine Wave Damping}
This test is adapted from \cite{skinnerOstriker_2013}.
Figure~\ref{fig:SineWaveDamping} shows the numerical result of sine wave damping test.
See \cite{chu_2018} for more information about the test setting.
SSP2332 as a second-order scheme display second order convergence rate.
PD-ARSs convergent in first order, as a price paid for realizability-preserving.
\begin{figure}[h]
  \centering
    \includegraphics[width=0.7\textwidth]{figures/SineWaveDamping}
   \caption{Relative error versus number of elements $N$ for the damping sine wave test. Results for different values of the absorption opacity $\sigma_{\Ab}$, employing various IMEX time stepping schemes, are compared.  Errors for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are SSP2332 ($+$), PD-ARS with SSPRK2 (circles) and PD-ARS with SSPRK3 (hexagram).  Black dash-dot reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively.}
  \label{fig:SineWaveDamping}
\end{figure}

\subsubsection{Sine Wave Diffusion}
The last test with known smooth solution is sine wave diffusion test.
It is adopted from \cite{radice_etal_2013}.
See \cite{chu_2018} for more information about the test setting.
Figure~\ref{fig:SineWaveDiffusionJ} shows the numerical result.
SSP2332 and PD-ARSs display third-order accuracy for the number density $\cJ$ and second-oder accuracy for $\cH_{x}$ and their errors are difficult to distinguish.
PD-ARS with SSP2332 behaviors as good as SSP2332 in the diffusion region but requires 1/3 less implicit solver.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionJ}
    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionH}
  \end{tabular}
   \caption{Absolute error for the number density $\cJ$ (left) and the number flux $\cH_{x}$ (right) versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are:  SSP2332 ($+$), PD-ARS with SSPRK2 (circles) and PD-ARS with SSPRK3 (hexagram). Black dash-dot on the left plot reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively. Black dash-dot on the right plot reference lines are proportional to $N^{-2}$ .}
   \label{fig:SineWaveDiffusionJ}
\end{figure}

### 2018-10-01 Documents/M1/Astronum_2018/sections/SpacialDiscretization.tex
Added conclusion section.

Divide the spatial domain $D$ into $N$ cells and denote the cell as $I_{i}$ with $i = 1,\ldots,N$, so that
\begin{equation*}
D = \cup_{i = 1}^{N} I_{i}.
\end{equation*}
Each of the cell has a volume/length of $\dx$.
Then we have cell average moment:
\begin{equation}
\bcM_{i} = \dfrac{1}{\dx} \int_{I_i}\bcM dx.
\end{equation}
Integrate Eq.~\eqref{eq:momentEquations} for each cell $I_{i}$ to have
\begin{equation}
\dfrac{d \bcM_{i}}{d t} = - \dfrac{1}{\dx} \left( \widehat{\bcF}(\bcM_{i},\bcM_{i+1}) -  \widehat{\bcF}(\bcM_{i-1},\bcM_{i})\right) + \f{1}{\tau}\,\cC(\bcM_{i}),
\end{equation}
where $\widehat{\bcF}(\vect{\cM}_{a},\vect{\cM}_{b})$ is numerical flux and $\f{1}{\tau}\,\cC(\bcM_{i})$ is the integrated collision term.
In this paper we use the simple Lax-Friedrichs flux that given by
\begin{equation}
  \widehat{\bcF}_{LF}(\vect{\cM}_{a},\vect{\cM}_{b})
  =\f{1}{2}\,\big(\,\bcF(\vect{\cM}_{a})+\bcF(\vect{\cM}_{b})-(\,\vect{\cM}_{b}-\vect{\cM}_{a}\,)\,\big).
\end{equation}
Therefore, by treating the flux term explicitly and the collision term implicitly, we have
\begin{align*}
\bcM_{i}^{n+1} = \widetilde{\bcM}^{n}_{i} + \f{\dt}{\tau}\,\cC(\bcM^{n+1}_{i})
\end{align*}
with
\begin{align*}
\widetilde{\bcM}^{n}_{i} 
& = \bcM_{i}^{n} - \dfrac{\dt}{\dx} \left( \widehat{\bcF}(\bcM^{n}_{i},\bcM^{n}_{i+1}) -  \widehat{\bcF}(\bcM^{n}_{i-1},\bcM^{n}_{i})\right)\nonumber \\
& = (1-\beta)\bcM_{i}^{n} + \beta\left[ \f{1}{2}\left( \bcM^{n}_{i+1}-\bcF(\bcM^{n}_{i+1})\right)  + \f{1}{2}\left( \bcM^{n}_{i-1}+\bcF(\bcM^{n}_{i-1})\right)\right]
\end{align*}
where $\beta := \dfrac{\dt}{\dx}$.
Assume $\bcM_{i}^{n}$ is realizable for all $i$. 
Lemma 3 in \cite{chu_2018} says $\bcM^{n+1}_{i}$ remains realizable given $\f{\dt}{\tau} > 0$ and $\widetilde{\bcM}^{n}_{i}$ is realizable.
$\widetilde{\bcM}^{n}_{i}$ is a convex combination of $\bcM_{i}^{n}$ and the bracket term if $\beta \in [0,1]$, so it is realizable if $\beta \in [0,1]$ and the bracket term is realizable.
And the realizability of the bracket term is given by Lemma 2 in \cite{chu_2018}.
To sum up, $\bcM^{n+1}_{i}$ remains realizable given $\beta \in [0,1]$.
It's the CFL condition of the Forward-Euler method employing on a transport only Boltzmann equation.

>>>>

Divide the spatial domain $D$ into $N$ cells and denote the cell as $I_{i}$ with $i = 1,\ldots,N$, so that
\begin{equation*}
D = \cup_{i = 1}^{N} I_{i}.
\end{equation*}
Each of the cell has a volume/length of $\dx$.
Then we have cell average moment:
\begin{equation}
\bcM_{i} = \dfrac{1}{\dx} \int_{I_i}\bcM dx.
\end{equation}
Integrate Eq.~\eqref{eq:momentEquations} for each cell $I_{i}$ to have
\begin{equation}
\dfrac{d \bcM_{i}}{d t} = - \dfrac{1}{\dx} \left( \widehat{\bcF}(\bcM_{i},\bcM_{i+1}) -  \widehat{\bcF}(\bcM_{i-1},\bcM_{i})\right) + \f{1}{\tau}\,\cC(\bcM_{i}),
\end{equation}
where $\widehat{\bcF}(\vect{\cM}_{a},\vect{\cM}_{b})$ is numerical flux and $\f{1}{\tau}\,\cC(\bcM_{i})$ is the integrated collision term.
In this paper we use the simple Lax-Friedrichs flux that given by
\begin{equation}
  \widehat{\bcF}_{LF}(\vect{\cM}_{a},\vect{\cM}_{b})
  =\f{1}{2}\,\big(\,\bcF(\vect{\cM}_{a})+\bcF(\vect{\cM}_{b})-(\,\vect{\cM}_{b}-\vect{\cM}_{a}\,)\,\big).
\end{equation}
Therefore, by treating the flux term explicitly and the collision term implicitly, we have
\begin{align*}
\bcM_{i}^{n+1} = \widetilde{\bcM}^{n}_{i} + \f{\dt}{\tau}\,\cC(\bcM^{n+1}_{i})
\end{align*}
with
\begin{align*}
\widetilde{\bcM}^{n}_{i} 
& = \bcM_{i}^{n} - \dfrac{\dt}{\dx} \left( \widehat{\bcF}(\bcM^{n}_{i},\bcM^{n}_{i+1}) -  \widehat{\bcF}(\bcM^{n}_{i-1},\bcM^{n}_{i})\right)\nonumber \\
& = (1-\beta)\bcM_{i}^{n} + \beta\left[ \f{1}{2}\left( \bcM^{n}_{i+1}-\bcF(\bcM^{n}_{i+1})\right)  + \f{1}{2}\left( \bcM^{n}_{i-1}+\bcF(\bcM^{n}_{i-1})\right)\right]
\end{align*}
where $\beta := \dfrac{\dt}{\dx}$.
Assume $\bcM_{i}^{n}$ is realizable for all $i$. 
Lemma 3 in \cite{chu_2018} says $\bcM^{n+1}_{i}$ remains realizable given $\f{\dt}{\tau} > 0$ and $\widetilde{\bcM}^{n}_{i}$ is realizable.
Substitute the Lax-Friedrichs flux formula in and $\widetilde{\bcM}^{n}_{i}$ becomes a convex combination of $\bcM_{i}^{n}$ and the square bracket term if $\beta \in [0,1]$.
So it is realizable if $\beta \in [0,1]$ and the square bracket term is realizable.
Given by Lemma 2 in \cite{chu_2018}, the square bracket term is realizable.
Therefore, $\bcM^{n+1}_{i}$ remains realizable given $\beta \in [0,1]$.
Note that $\beta \in [0,1]$ is also the CFL condition of the Forward-Euler method employing on a transport only Boltzmann equation.

### 2018-10-01 Documents/M1/Astronum_2018/sections/TwoMoments.tex
Added conclusion section.

The inequalities in Eq.\ref{eq:eddingtonFactorBounds} desire more attention than have be given.
They have the equal importance as the inequalities in Eq.\ref{eq:MomentsBounds} in preserving the validity of the simulation result.
However, as Fig.~\eqref{fig:EddingtonFactorsWithDifferentClosure} shows, not all the algebraic closures satisfy the Eddington factor bounds Eq.\ref{eq:eddingtonFactorBounds}: Kershaw\cite{kershaw_1976}, Wilson [citation], Levermore\cite{levermore_1984}, Minerbo \cite{minerbo_1978}, Janka 2 [citation] may work fine with neutrino low occupied setting, but not for high occupancy situation; Janka 1 [citation] can be dangerous for both the cases. The only algebraic closure that remains in the bounds among those seven plotted closures is Cernohorsky \& Bludman's \cite{cernohorskyBludman_1994}.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/Closures0_10}
    \includegraphics[width=0.5\textwidth]{figures/Closures0_90}
  \end{tabular}
   \caption{Plot of Eddington factors $\chi$ versus flux factor $h$ for different values of $\cJ$ for various algebraic closures: $\cJ=0.1$ (left panel, low occupied) and $\cJ=0.9$ (right panel, high occupied).  In each panel we plot the Eddington factors of Kershaw (red), Wilson (yellow), Levermore (green), Minerbo (light blue), Cernohorsky \& Bludman (blue) and Janka (purple and pink) closures.  We also plot $\chi_{\mbox{\tiny min}}$ and $\chi_{\mbox{\tiny max}}$ defined in Eq.~\eqref{eq:eddingtonFactorBounds} (lower and upper dash black lines, respectively).}
  \label{fig:EddingtonFactorsWithDifferentClosure}
\end{figure}

>>>>

The inequalities in Eq.\ref{eq:eddingtonFactorBounds} desire more attention than have be given.
They have the equal importance as the inequalities in Eq.\ref{eq:MomentsBounds} in preserving the validity of the simulation result.
However, as Fig.~\eqref{fig:EddingtonFactorsWithDifferentClosure} shows, not all the algebraic closures satisfy the Eddington factor bounds Eq.\ref{eq:eddingtonFactorBounds}: Kershaw\cite{kershaw_1976}, Wilson\cite{wilson1975,leblanc1970}, Levermore\cite{levermore_1984}, Minerbo \cite{minerbo_1978}, Janka 2\cite{janka1992} may work fine with neutrino low occupied setting, but not for high occupancy situation; Janka 1\cite{janka1991} can be dangerous for both the cases. The only algebraic closure that remains in the bounds among those seven plotted closures is Cernohorsky \& Bludman's \cite{cernohorskyBludman_1994}.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/Closures0_10}
    \includegraphics[width=0.5\textwidth]{figures/Closures0_90}
  \end{tabular}
   \caption{Plot of Eddington factors $\chi$ versus flux factor $h$ for different values of $\cJ$ for various algebraic closures: $\cJ=0.1$ (left panel, low occupied) and $\cJ=0.9$ (right panel, high occupied).  In each panel we plot the Eddington factors of Kershaw (red), Wilson (yellow), Levermore (green), Minerbo (light blue), Cernohorsky \& Bludman (blue) and Janka (purple and pink) closures.  We also plot $\chi_{\mbox{\tiny min}}$ and $\chi_{\mbox{\tiny max}}$ defined in Eq.~\eqref{eq:eddingtonFactorBounds} (lower and upper dash black lines, respectively).}
  \label{fig:EddingtonFactorsWithDifferentClosure}
\end{figure}
